Introducing the JuliaHealth Observational Health Research Subecosystem

### 2023-12-30 paper.tex
Added some figure captions, enumerations, and formatting options

@@

% TODO::EASY Reference code repository
Using the JuliaHealth tools developed, we enumerate an example observational health study workflow that would be used in practice.\footnotemark
% EXPLAIN::EASY In enumeration talk about what Eunomia is 
% EXPLAIN::EASY In enumeration discuss MIMIC III

For this paper, we used two datasets: a synthetic patient dataset called \textit{Eunomia} and the MIMIC III patient dataset which had been converted into the OMOP CDM to illustrate the workflow provided by the Observational Health sub-ecosystem.\cite{physionet, hripcsakCharacterizingTreatmentPathways2016, johnson2016mimic, johnson2018mimic}

@@

\item \textbf{Eunomia} -

\item \textbf{MIMIC III} -

@@

% EXPLAIN::EASY What is this doing here?
% TODO::HARD Update DBConnector
% TODO::MEDIUM Maybe remove package imports and mention them within the caption instead

\begin{listing}[!ht]
\begin{minted}[breaklines,escapeinside=||,mathescape=true, numbersep=3pt, gobble=2, frame=lines, fontsize=\scriptsize, framesep=2mm]{julia}
import DBConnector: 
  DBConnection
import HealthSampleData: 
  Eunomia
import OMOPCDMCohortCreator as OCC

@@

occ.GenerateTables(conn)
\end{minted}
\caption{This illustrates how to connect to the Eunomia database}
\label{listing:connection}
\end{listing}

@@

% TODO::HARD Maybe remove package imports and mention them within the caption instead
\begin{listing}[!ht]
\begin{minted}[breaklines,escapeinside=||,mathescape=true, numbersep=3pt, gobble=2, frame=lines, fontsize=\scriptsize, framesep=2mm]{julia}
using JSON3 
using OHDSICohortExpressions

import DBInterface as DBI

cohort = JSON3.read("strep_throat.json")
cohort_expression = cohort[:items][1][:expression]

model = Model(cdm_version = v"5.3.1", 
              cdm_schema = "main",
              vocabulary_schema = "main", 
              results_schema = "main",
              target_schema = "main", 
              target_table = "cohort");

sql = translate(cohort_expression, 
                dialect = :sqlite, 
                model = model, 
                cohort_definition_id = 1);

for query in split(sql, ";")[1:end-1]
  DBI.execute(conn, query)
end
\end{minted}
% TODO::EASY Add a caption
\caption{}
% TODO::EASY Add a label
\label{listing:}
\end{listing}

@@

% TODO::EASY Maybe remove package imports and mention them within the caption instead
\begin{listing}[!ht]
\begin{minted}[breaklines,escapeinside=||,mathescape=true, numbersep=3pt, gobble=2, frame=lines, fontsize=\scriptsize, framesep=2mm]{julia}
C = GetCohortSubjects(1, conn)
rename!(C, :subject_id => :person_id)

C_race = occ.GetPatientRace(C.person_id, conn)
C_gender = occ.GetPatientGender(C.person_id, conn)

C_age_group = occ.GetPatientAgeGroup(
  C.person_id, 
  conn; 
)
\end{minted}
% TODO::EASY Add a caption
\caption{}
% TODO::EASY Add a label
\label{listing:}
\end{listing}

@@

% TODO::EASY Maybe remove package imports and mention them within the caption instead

\begin{listing}[!ht]
\begin{minted}[breaklines,escapeinside=||,mathescape=true, numbersep=3pt, gobble=2, frame=lines, fontsize=\scriptsize, framesep=2mm]{julia}
import DataFrames as DF

C_features = DF.outerjoin(C_race,
                          C_gender,
                          C_age_group;
                          on = :person_id, 
                          matchmissing = :equal)

C_features = C_features[:, DF.Not(:person_id)]

C_groups = DF.groupby(C_features, 
                      [
                        :race_concept_id, 
                        :gender_concept_id, 
                        :age_group
                      ])

C_groups = DF.combine(C_groups, DF.nrow => :count)
\end{minted}
% TODO::EASY Add a caption
\caption{}
% TODO::EASY Add a label
\label{listing:}
\end{listing}

@@

% EXPLAIN::EASY Move this to the caption of the table
Additionally, throughout the process, one can perform relevant analysis methods.
Here, we'll conduct a crude prevalence calculation across the patient cohort.

@@

\begin{table}[!ht]
    \centering
    \begin{tabular}{|l|l|l|l|l|l|}
    \hline
        Race & Gender & Age Group & Cohort & Total & Prev. (\%) \\ \hline
        B/AA & F & 60 - 64 & 13 & 21 & 61.90 \\ \hline
        W & F & 50 - 54 & 59 & 111 & 53.15 \\ \hline
        W & F & 55 - 59 & 62 & 97 & 63.91 \\ \hline
        A & M & 80 - 84 & 6 & 10 & 60.00 \\ \hline
        UNK & F & 60 - 64 & 29 & 39 & 74.35 \\ \hline
        B/AA & M & 65 - 69 & 12 & 18 & 66.66 \\ \hline
    \end{tabular}
    % TODO::EASY Add a caption
    \caption{}
    % TODO::EASY Add a label
    \label{table:}
\end{table}

@@

% TODO::MEDIUM Can this be simplified?
\begin{listing}[!ht]
\begin{minted}[breaklines,escapeinside=||,mathescape=true, numbersep=3pt, gobble=2, frame=lines, fontsize=\scriptsize, framesep=2mm]{julia}
CharacterizePatients(patients) = @chain patients begin
    _[:, DF.Not(:person_id)]
    DF.groupby(_, names(_))
    DF.combine(_, DF.nrow => :count)
    occ.ExecuteAudit
end
\end{minted}
% TODO::EASY Add a caption
\caption{}
% TODO::EASY Add a label
\label{listing:}
\end{listing}

>>>>

\usepackage[labelfont=bf]{caption}
\captionsetup[table]{skip=4pt}
\captionsetup{font=scriptsize}

@@

% TODO::EASY Reference code repository
Using the JuliaHealth tools developed, we enumerate an example observational health study workflow that would be used in practice.\footnotemark

For this paper, we used two datasets:

@@

% TODO::EASY Cite https://github.com/synthetichealth/synthea
\item \textbf{Eunomia} - this is a synthetic patient dataset created from the open source synthetic patient generator, \textit{Synthea}, and hosted within the HADES and JuliaHealth ecosystems.

\item \textbf{MIMIC III} - a deidentified datasets consisting of inpatient visits from Beth Israel Deaconness Hospital in Boston, Massachussetts.
The particular version of the dataset that we used was transformed into the OMOP CDM using tools from MIT LCP and PhysioNet. \cite{physionet, hripcsakCharacterizingTreatmentPathways2016, johnson2016mimic, johnson2018mimic}

@@

% EXPLAIN::EASY What is this doing here?
% TODO::HARD Update DBConnector
\begin{listing}[!ht]
\begin{minted}[breaklines,escapeinside=||,mathescape=true, numbersep=3pt, gobble=2, frame=lines, fontsize=\scriptsize, framesep=2mm]{julia}
import DBConnector: 
  DBConnection
import HealthSampleData: 
  Eunomia

@@

occ.GenerateTables(conn)
\end{minted}
\caption{\textbf{Configuring OMOP CDM Database Connection.} Using \textit{DBConnector.jl}, connection to the \textit{Eunomia} SQLite database can be made. Then, \textit{OMOPCDMCohortCreator.jl} (occ) is used to generate connection details used internally by the package for the rest of the session.}
\label{listing:connection}
\end{listing}

@@

% TODO::HARD Maybe remove package imports and mention them within the caption instead
\begin{listing}[!ht]
\begin{minted}[breaklines,escapeinside=||,mathescape=true, numbersep=3pt, gobble=2, frame=lines, fontsize=\scriptsize, framesep=2mm]{julia}
model = oce.Model(cdm_version = v"5.3.1", 
              cdm_schema = "main",
              vocabulary_schema = "main", 
              results_schema = "main",
              target_schema = "main", 
              target_table = "cohort");

sql = oce.translate(cohort_expression, 
                dialect = :sqlite, 
                model = model, 
                cohort_definition_id = 1);
\end{minted}
\caption{\textbf{Creating a Cohort Using a Computable Phenotype Definition.} \textit{OHDSICohortExpressions.jl} (oce) ingests a computable phenotype definition to generate a cohort of patients. The JSON is translated into an intermediate SQL Syntax Tree, instantiated in SQL (SQLite) targeting an OMOP CDM v5.3.1 database, and then populates the database's cohort table.}
\label{listing:cohort_creation}
\end{listing}

@@

\begin{listing}[!ht]
\begin{minted}[breaklines,escapeinside=||,mathescape=true, numbersep=3pt, gobble=2, frame=lines, fontsize=\scriptsize, framesep=2mm]{julia}
patients = occ.GetCohortSubjects(1, conn)

patients_race = occ.GetPatientRace(C.person_id, conn)
patients_gender = occ.GetPatientGender(C.person_id, conn)

patients_age_group = occ.GetPatientAgeGroup(
  C.person_id, 
  conn; 
)
\end{minted}
\caption{\textbf{Find Demographic Characteristics of Cohort.} Demographic properties of a cohort can be quickly queried from a database.}
\label{listing:demographics}
\end{listing}

@@

\begin{listing}[!ht]
\begin{minted}[breaklines,escapeinside=||,mathescape=true, numbersep=3pt, gobble=2, frame=lines, fontsize=\scriptsize, framesep=2mm]{julia}
strep_df = df.combine(patient_groups, df.nrow => :count)

# Creator code to process data and generate counts...

strep_df.prev = strep_df.count ./ strep_df.total_count 
\end{minted}
\caption{\textbf{Calculating crude prevalence rates.} Once demographic information is extracted, processing can be performed using tools such as \textit{DataFrames.jl} (df) to calculate metrics such as prevalence rates in a very straightforward manner.}
\label{listing:prevalence}
\end{listing}

@@

@@

\begin{table}[!ht]
    \centering
    \begin{tabular}{|l|l|l|l|l|l|}
    \hline
        Race & Gender & Age Group & Cohort & Total & Prev. (\%) \\ \hline
        B/AA & F & 60 - 64 & 13 & 21 & 61.90 \\ \hline
        W & F & 50 - 54 & 59 & 111 & 53.15 \\ \hline
        W & F & 55 - 59 & 62 & 97 & 63.91 \\ \hline
        A & M & 80 - 84 & 6 & 10 & 60.00 \\ \hline
        UNK & F & 60 - 64 & 29 & 39 & 74.35 \\ \hline
        B/AA & M & 65 - 69 & 12 & 18 & 66.66 \\ \hline
    \end{tabular}
    % TODO::EASY Add a caption
    \caption{\textbf{Example of a Strep Throat Prevalence Calculation.} This illustrates the sort of results one can quickly generate using the OHS. Note, that although the prevalence values in this table are based on synthetic data and are not realistic, the output looks very similar to what one could expect to generate.}
    \label{table:eunomia_prevalence}
\end{table}

@@

% TODO::MEDIUM Can this be simplified?
\begin{listing}[!ht]
\begin{minted}[breaklines,escapeinside=||,mathescape=true, numbersep=3pt, gobble=2, frame=lines, fontsize=\scriptsize, framesep=2mm]{julia}
CharacterizePatients(patients) = @chain patients begin
    _[:, df.Not(:person_id)]
    df.groupby(_, names(_))
    df.combine(_, df.nrow => :count)
    occ.ExecuteAudit
end
\end{minted}
% TODO::EASY Add a caption
\caption{}
% TODO::EASY Add a label
\label{listing:}
\end{listing}
